Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-05-23 manuscript_results.Rmd
update

# Heatmap plot with models ordered by WeekAhead == 1  
plot_m_heatmap <- ggplot(p_df, aes(x = factor(WeekAhead), y = Model, fill = m)) +
  geom_tile(aes(fill = ifelse(Significance == "Significant", m, NA)), color = "black") +  # Fill only significant ones
  geom_tile(data = p_df %>% filter(Significance == "Not Significant"), 
            aes(x = factor(WeekAhead), y = Model), fill = "gray80", color = "black") +  # Gray for non-significant
  geom_text(aes(label = paste0("m:", round(m, 1),", sd:", round(sd, 1))), 
            color = "black", size = 4) +  # Add text labels
  scale_fill_gradient(low = "cyan", high = "red", na.value = "gray80", 
                      name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
  labs(title = "Mean differences between models' WIS and the baseline across 48 states",
       subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
         caption = "m = mean difference (%), sd = standard deviation",
       x = "Target Ahead",
       y = "") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  plot.caption = element_text(size = 16),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),) +
  facet_wrap(~ Model_Type, scales = "free_y")  # Separate NoLg and other models

@@

@@

ggsave(reg2, height = 6, width = 8, filename="Fig7.jpg")  
#ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
```

@@

ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("1 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Percentage (%)",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("2 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Percentage (%)",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("3 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill =  "Percentage (%)",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

```{r}
ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("4 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Percentage (%)",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

@@

reg2
ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
```

@@

# Plot the difference using ggplot2
model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
  theme_minimal() +
  labs(
    title = "Mean percentage of WIS differences across all states",
    subtitle = "AUTO_AVG_LB compared to baseline",
    x = "",
    y = "Difference (%)"
  ) +
  scale_y_continuous(
    limits = c(-100, 100),
    breaks = seq(-100, 100, by = 20)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  scale_color_manual(
    values = c("Week 1" = "#ffcccc",
               "Week 2" = "#ff6666",
               "Week 3" = "#cc0000",
               "Week 4" = "#990000")
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

ggsave(model_improv, height = 4.5, width = 9,  filename="Fig10.jpg")  
```

>>>>

# Heatmap plot with models ordered by WeekAhead == 1  
plot_m_heatmap <- ggplot(p_df, aes(x = factor(WeekAhead), y = Model, fill = m)) +
  geom_tile(aes(fill = ifelse(Significance == "Significant", m, NA)), color = "black") +  # Fill only significant ones
  geom_tile(data = p_df %>% filter(Significance == "Not Significant"), 
            aes(x = factor(WeekAhead), y = Model), fill = "gray80", color = "black") +  # Gray for non-significant
  geom_text(aes(label = paste0("m:", round(m, 1),", sd:", round(sd, 1))), 
            color = "black", size = 4) +  # Add text labels
  scale_fill_gradient2(
  low = "cyan", 
  mid = "white", 
  high = "red", 
  midpoint = 0,     # <-- Sets 0 as the midpoint
  na.value = "gray80", 
  name = "Mean difference", 
  limits = c(-35, 35)       
)+

  labs(title = "Mean differences between models' WIS and the baseline across 48 states",
       subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
         caption = "m = mean difference, sd = standard deviation",
       x = "Target Ahead",
       y = "") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  plot.caption = element_text(size = 16),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),) +
  facet_wrap(~ Model_Type, scales = "free_y")  # Separate NoLg and other models

@@

regression_models

best regression plot

```{r}
plot_list <- list()  # store plots

for(i in 1:4) {

  # Combine WIS data and population
  WIS_pop_data <- inner_join(WIS_dataframes[[i]], pop_data, by = "STATE")

  # Fit linear model
  model <- lm(AUTO_AVG ~ Resident_population_2020, data = WIS_pop_data)
  model_summary <- summary(model)

  # Save R² and p-values
  r_squared <- round(model_summary$r.squared, 3)
  p_value <- signif(model_summary$coefficients[2, 4], 3)

  # Create plot
  p <- ggplot(WIS_pop_data, aes(x = Resident_population_2020, y = AUTO_AVG)) +
    geom_point(color = "steelblue", size = 2) +
    geom_smooth(method = "lm", se = FALSE, color = "darkred", linetype = "dashed") +
    labs(
      title = paste("Regression","Target Week", i),
      subtitle = paste("R² =", r_squared, "| p =", p_value),
      x = "Resident Population (2020)",
      y = "AUTO_AVG"
    ) +
    theme_minimal()

  # Store the plot
  plot_list[[i]] <- p
}

# Combine the plots vertically
combined_regressions <- wrap_plots(plot_list, ncol = 1)

# Display (optional)
combined_regressions

# Save the combined plot
#ggsave("Figx.png", combined_plot, width = 3, height = 5, dpi = 600)
```

@@

# Combine the plots vertically
reg2_scatter <- reg2 | combined_regressions 
reg2_scatter
ggsave(reg2_scatter, height = 6, width = 12, filename="Fig7.jpg")  
#ggsave(reg2, filename="reg2.jpg",height = 8, width = 10)
```

@@

ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("1 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("2 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("3 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill =  "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

```{r}
ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("4 weeks ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "Mean difference",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

best regression plot

# Prepare a list to hold your plots
bric_plots <- list()

for (i in 1:4) {
  # 1) Join the WIS and BRIC data
  WIS_bric <- inner_join(
    WIS_dataframes[[i]],
    bric_by_state,
    by = "STATE"
  )

  # 2) Fit the linear model
  model <- lm(
    percentage_improvement ~ weighted_mean_z_bric.infrastructure,
    data = WIS_bric
  )
  ms <- summary(model)

  # 3) Extract R² and p‐value
  r2 <- round(ms$r.squared, 3)
  p  <- signif(ms$coefficients[2, 4], 3)

  # 4) Build the ggplot
  p <- ggplot(WIS_bric,
              aes(x = weighted_mean_z_bric.infrastructure,
                  y = percentage_improvement)) +
    geom_point(color = "steelblue", size = 2) +
    geom_smooth(method = "lm", se = FALSE, color = "darkred") +
    labs(
      title    = paste("Regression","Target Week", i),
      subtitle = paste0("R² = ", r2, "   p = ", p),
      x        = "BRIC infrastructure (2015)",
      y        = "Difference"
    ) +
    theme_minimal()

  # 5) Store it
  bric_plots[[i]] <- p
}

# Combine the plots vertically
combined_regressions <- wrap_plots(bric_plots, ncol = 1)

# Display 
combined_regressions

@@

reg2_scatter <- reg2 | combined_regressions 
reg2_scatter
ggsave(reg2_scatter, height = 6, width = 12, filename="Fig9.jpg")  
#ggsave(reg2, filename="Fig9.jpg",height = 6, width = 8)
```

@@

# Plot the difference using ggplot2
model_improv<-ggplot(summed_data_all, aes(x = Julian_date, y = (Difference), color = Week)) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed", size = 0.8) +
  theme_minimal() +
  labs(
    title = "Weekly mean WIS differences across all states",
    subtitle = "AUTO_AVG_LB compared to baseline",
    x = "",
    y = "Mean difference"
  ) +
  scale_y_continuous(
    limits = c(-100, 100),
    breaks = seq(-100, 100, by = 20)
  ) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  scale_color_manual(
    values = c(
    "Week 1" = "#00BFC4",  # Bright cyan
    "Week 2" = "#F8766D",  # Bright coral red
    "Week 3" = "#7CAE00",  # Lime green
    "Week 4" = "#C77CFF"   # Bright purple
  )
  ) +
  theme(
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

@@

ggsave(model_improv, height = 4.5, width = 9,  filename="Fig10.jpg")  
```
hospitalizations

total_hospitalizations_by_week <- AUTO_ADJACENT_WEEK1 %>%
  filter(epiweek >= 40 | epiweek <= 20) %>%  # Apply filter AFTER epiweek exists
  group_by(target_end_date) %>%
  summarise(mean_cases = sum(cases, na.rm = TRUE))

hospital_plot <-ggplot(total_hospitalizations_by_week, aes(x = target_end_date, y = mean_cases)) +
  geom_bar(stat = "identity",fill = "steelblue") +
  labs(title = "Total influenza hospitalizations on the 48 states",
       x = "",
       y = "Influenza hospitalizations") +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b %y"
  ) +
  theme_minimal()+
  theme(plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1))

hospital_plot

# Combine the plots vertically
combined_plot <- hospital_plot / model_improv

# Display (optional)
combined_plot

# Save the combined plot
ggsave("Fig10.png", combined_plot, width = 7, height = 7, dpi = 600)
